Unknown

### 2025-06-28 ms/ms.qmd

<!-- Paragraph 7 (summarize secondary results) -->
The pattern of AA across wild tomatoes strongly supports the plasticity hypothesis, argues against the acclimatory hypothesis, and provides only weak support for the constitutive hypothesis. Plastic changes in leaf thickness and/or packing density, summarized by the bulk leaf mass per area (LMA), likely mediate the effect of growth light intensity on AA. LMA increased in sun grown plants in all accessions by an average of X% [CI], similar to plastic responses in many species [@poorter_metaanalysis_2019]. When LMA is included in the model, the effect of growth light intensity on AA is no longer significant based on model comparison (Table ? - need to check out what the interaction with measurement light intensity means) and LMA is weakly positively correlated with AA (Figure X). In contrast, LMA did not mediate indirect effects of native PAI on AA (Table ?). Many anatomical traits underlie LMA [@john_anatomical_2017] and future research will be needed identify which particular traits, such as leaf thickness or mesophyll porosity, are responsible for mediating AA.

@@

@@

Problems to bring up in discussion:
- all our species were amphi, so they may all live in PAI below threshold to favor amphi
- other microclimate diffs not considered
- don't know specific anatomical change

@@

We fit Bayesian models with MCMC sampling in the probabilistic programming language *Stan* [@stan_development_team_stan_2025] using the *R* package **brms** version `r fit_aa$version$brms` [@burkner_brms_2017]. We used CmdStan version `r fit_aa$version$cmdstan` and **cmdstanr** version `r fit_aa$version$cmdstanr` [@gabry_cmdstanr_2025] to interface with *R* version `r getRversion()` [@r_core_team_r:_2025]. We sampled the posterior distribution from `r dim_fit_aa[2]` chains with `r dim_fit_aa[1]` iterations each after `r dim_fit_aa[1]` warmup iterations per chain. We estimated parameters and confidence intervals as the median and 95% quantile intervals of the posterior, respectively. We chose the number of chains, warmup and sampling iterations, and maximum treedepth so that parameter estimates converged ($\hat{R} < 1.01$ [@gelman_inference_1992]) and the effective sample size (ESS) for each parameter was $> 10^3$.

>>>>

<!-- Paragraph 7 (summarize secondary results) -->
The pattern of AA across wild tomatoes strongly supports the plasticity hypothesis, argues against the acclimatory hypothesis, and provides only weak support for the constitutive hypothesis. Plastic changes in leaf thickness and/or packing density, summarized by the bulk leaf mass per area (LMA), likely mediate the effect of growth light intensity on AA. LMA increased in sun grown plants in all accessions by an average of X% [CI], similar to plastic responses in many species [@poorter_metaanalysis_2019]. When LMA is included in the model, the effect of growth light intensity on AA is no longer significant based on model comparison (Table ? - need to check out what the interaction with measurement light intensity means) and LMA is weakly positively correlated with AA (Figure X). In contrast, LMA did not mediate indirect effects of native PAI on AA (Table ?). Many anatomical traits underlie LMA [@john_anatomical_2017] and future research will be needed identify which particular traits, such as leaf thickness or mesophyll porosity, are responsible for mediating AA. Our study is also limited by the fact that we could not directly measure the stomatal conductance ratio and our method for estimating AA requires large sample sizes to overcome several sources of measurement error (see supporting information). Future experiments measuring AA with a dual sided chamber [@wall; @marquez] can overcome these limitations.

@@

big implications - if trees were amphi, more more carbon sink or reduced transpiration

### MANUSCRAPS
Problems to bring up in discussion:
- all our species were amphi, so they may all live in PAI below threshold to favor amphi
- other microclimate diffs not considered
- don't know specific anatomical change

@@

@@

We fit Bayesian models with MCMC sampling in the probabilistic programming language *Stan* [@stan_development_team_stan_2025] using the *R* package **brms** version `r fit_aa$version$brms` [@burkner_brms_2017]. We used CmdStan version `r fit_aa$version$cmdstan` and **cmdstanr** version `r fit_aa$version$cmdstanr` [@gabry_cmdstanr_2025] to interface with *R* version `r getRversion()` [@r_core_team_r:_2025]. We sampled the posterior distribution from `r dim_fit_aa[2]` chains with `r dim_fit_aa[1]` iterations each after `r dim_fit_aa[1]` warmup iterations per chain. We estimated parameters and confidence intervals as the median and 95% quantile intervals of the posterior, respectively. We chose the number of chains, warmup and sampling iterations, and maximum treedepth so that parameter estimates converged ($\hat{R} < `r aa_args['max_rhat']`$ [@gelman_inference_1992]) and the effective sample size (ESS) for each parameter was $> `r aa_args['min_ess']`$.
